---
title: "Top Streamed Songs of 2017 - 2019"
output: html_document
---
Group: 313A
 
Names: Danny Soulas, Laura Dinh, Torrey Coan, Nicholas Tam 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning = FALSE,
                      cache=TRUE, autodep=TRUE, cache.comments=FALSE)
library(tidyverse)
library(scales)
library(modelr)
library(tidymodels)
library(kableExtra)
source("scripts/viridis.R")
source("scripts/ggprob.R")
```


```{r, echo=FALSE, include=FALSE}
songs_2017 = read_csv("spotify-2017.csv") %>% mutate(charting_yr = 2017, rank = 1:100) 
songs_2018 = read_csv("spotify-2018.csv") %>% mutate(charting_yr = 2018, rank = 1:100)
songs_2019 = read_csv("spotify-2019.csv") %>% mutate(charting_yr = 2019, rank = 1:50)

all_songs = rbind(songs_2017,songs_2018,songs_2019)
```

## Introduction

Music is very personal. Music tastes change from person to person. Our feelings and experiences can affect the way we react to a song. Despite our different tastes, there are songs that are able to appeal to a wider audience. These songs can rack up over a billion of streams. But what makes these songs so appealing? Do specific traits affect a song's overall stream-ability? For our project, we decided to take a look at two specific traits of songs: tempo and happiness, and examine how they changed in recent years. More specifically, we were interested in these questions:

- Did the tempo of top streamed songs get faster from 2017 to 2019?

  - Are tempo and rank correlated?

- Did the top streamed songs get happier from 2017 to 2019?

  - Are tempo and valence correlated?

Although it may seem songs have gotten faster and happier in the past few years, we did not find any statistically significant evidence that the tempo and happiness of songs have increased from 2017 to 2019.

## Background

Spotify is one of the most widely used streaming apps in the world with over 381 million users. The platform has been able to collect data from users and produce playlists containing top streamed songs. We used these top streamed song playlists generated by Spotify to examine our questions.


### About the Data

The Spotify Top Streamed 2017, 2018, and 2019 sets used were created by Nadin Tamer. The data was created by using [the Organize Your Music webite](http://organizeyourmusic.playlistmachinery.com/)^[Organize Your Music: (http://organizeyourmusic.playlistmachinery.com/)], a web app that allows users to break down different characteristics of their music consumption. It intakes a Spotify playlist and returns scores of certain variables for each song. To create the top streamed songs, playlists containing the top streamed songs from 2017, 2018, and 2019 were out into the engine. For each year's data set, each row is one song, each ordered from highest ranked to lowest rank. 

The data sets were uploaded to Kaggle by Nadin Tamer. Here are the links to the three data sets:

- [The Top Streamed Songs for 2017](https://www.kaggle.com/nadintamer/top-tracks-of-2017)^[Dataset for 2017: (https://www.kaggle.com/nadintamer/top-tracks-of-2017)]
- [The Top Streamed Songs for 2018](https://www.kaggle.com/nadintamer/top-spotify-tracks-of-2018)^[Dataset for 2018: (https://www.kaggle.com/nadintamer/top-spotify-tracks-of-2018)]
- [The Top Streamed Songs for 2019](https://www.kaggle.com/nadintamer/top-spotify-tracks-of-2019)^[Dataset for 2019: (https://www.kaggle.com/nadintamer/top-spotify-tracks-of-2019)]

From these data sets, we specifically examined tempo, valence, and rank. Tempo is the speed or pace of a given song. It is measured in beats per minute, but it can also be described with tempo indicators. Here are a few tempo indicators that demonstrate the typical tempo range of songs:

- Grave – slow and solemn (20–40 BPM)
- Lento – slowly (40–45 BPM)
- Largo – broadly (45–50 BPM)
- Adagio – slow and stately (literally, “at ease”) (55–65 BPM)
- Adagietto – rather slow (65–69 BPM)
- Andante – at a walking pace (73–77 BPM)
- Moderato – moderately (86–97 BPM)
- Allegretto – moderately fast (98–109 BPM)
- Allegro – fast, quickly and bright (109–132 BPM)
- Vivace – lively and fast (132–140 BPM)
- Presto – extremely fast (168–177 BPM)
- Prestissimo – even faster than Presto (178 BPM and over)

For more about tempo indicators [click here](https://symphonynovascotia.ca/faqs/symphony-101/how-do-musicians-know-how-fast-to-play-a-piece-and-why-are-the-terms-in-italian/)^[Tempo Definitions: (https://symphonynovascotia.ca/faqs/symphony-101/how-do-musicians-know-how-fast-to-play-a-piece-and-why-are-the-terms-in-italian/)]

Valence is specific to Spotify's API. It describes the musical positiveness conveyed by a track and is measure on a scale of 0 to 1. Tracks with a higher valence sound more positive while tracks with lower valence sound more sad. 

For more information about valence, here are links that provide more information: 

- [Valence Definition](https://www.theverge.com/tldr/2018/2/5/16974194/spotify-recommendation-algorithm-playlist-hack-nelson)^[Term Definitions: (https://www.theverge.com/tldr/2018/2/5/16974194/spotify-recommendation-algorithm-playlist-hack-nelson)]
- [How a Computer Determines Valence](https://web.archive.org/web/20170422195736/http://blog.echonest.com/post/66097438564/plotting-musics-emotional-valence-1950-2013)^[Where valence comes from: (https://web.archive.org/web/20170422195736/http://blog.echonest.com/post/66097438564/plotting-musics-emotional-valence-1950-2013)]

Rank represents how many streams a song has. A rank of 1 means the song was the top streamed song for the given year, a rank of 2 means the song was the second top streamed song for the given year, and so on.

#### Variables from Data

```{r, echo=FALSE}
song_variables = tibble(
  Name = c("name", "artists","valence", "tempo", "rank"),
  Description = c("A string of the song name",
                  "A string of the artist(s)' name(s)",
                  "A real number between 0 and 1 that represents the happiness of a song",
                  "A real number greater than or equal to 0 that represents the speed at which a passage of music is played. It is measured in beats per minute (BPM)",
                  "The popularity of the song for the year"))

song_variables %>% 
  kable(caption = "Key Variables from the Spotify Top Streamed Songs Data") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```


## Analysis

### Did the top streamed songs get faster from 2017 to 2019?

First, we summarize the tempo data for all three years. 

```{r, echo=FALSE}
tempo_2017 = songs_2017 %>% 
  select(tempo) %>% 
  summarize(n = n(),mean = mean(tempo), sd = sd(tempo), max = max(tempo), min = min(tempo))

tempo_2018 = songs_2018 %>% 
  select(tempo) %>% 
  summarize(n = n(),mean = mean(tempo), sd = sd(tempo), max = max(tempo), min = min(tempo))

tempo_2019 = songs_2019 %>% 
  select(tempo) %>% 
  summarize(n = n(),mean = mean(tempo), sd = sd(tempo), max = max(tempo), min = min(tempo))

tempo = bind_rows(tempo_2017, tempo_2018, tempo_2019)

years = tibble(
  year = c(2017, 2018, 2019)
)

bind_cols(years, tempo) %>% 
  kable(caption = "Tempo Summary") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```

- $year$ is the year the songs are from
- $n$ is the number of songs
- $mean$ is the mean tempo for the given year
- $sd$ is the standard deviation of the songs' tempo
- $max$ is the song with the fastest tempo for the year
- $min$ is the song with the slowest tempo for the year

From the table, we can see that the mean increases from between all three years, but is this increase statistically significant? 

#### T Tests

To test this, we ran an unpaired T Test (Two Mean Test) of tempos between 2017 and 2018:


```{r, echo=FALSE}
# combine tempo for 2017 and 2019
tempo_data_17 = songs_2017 %>% 
  select(tempo, rank) %>% 
  mutate(year = "2017")

tempo_data_18 = songs_2018 %>% 
  select(tempo, rank) %>% 
  mutate(year = "2018")

tempo_data = bind_rows(tempo_data_17, tempo_data_18)

tempo_wide = tempo_data %>% 
  pivot_wider(names_from = year, values_from = tempo)

head(tempo_wide, 5) %>% 
  kable(caption = "Tempo between 2017 and 2018 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```



```{r, echo=FALSE}
ggplot(tempo_data, aes(x = year, y = tempo, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Tempo") +
  ggtitle("Boxplots of Tempos in the Top Streamed Songs of 2017 and 2018") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2017 and the other for 2018

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = tempo_data %>% 
  filter(year == "2017") %>% 
  pull(tempo)

y = tempo_data %>% 
  filter(year == "2018") %>% 
  pull(tempo)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2017 = stats$estimate[1],
  mean_2018 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean tempo for the top streamed songs of 2017 is different than the mean tempo for the top streamed songs of 2018 (p = 0.431, unpaired t-test, df = 197.825)

Let's run the same unpaired t-test for 2018 and 2019:


```{r, echo=FALSE}

tempo_data_19 = songs_2019 %>% 
  select(tempo, rank) %>% 
  mutate(year = "2019")

tempo_data = bind_rows(tempo_data_18, tempo_data_19)

tempo_wide = tempo_data %>% 
  pivot_wider(names_from = year, values_from = tempo)

head(tempo_wide, 5) %>% 
  kable(caption = "Tempo between 2018 and 2019 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```



```{r, echo=FALSE}
ggplot(tempo_data, aes(x = year, y = tempo, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Tempo") +
  ggtitle("Boxplots of Tempos in the Top Streamed Songs of 2018 and 2019") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2018 and the other for 2019

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = tempo_data %>% 
  filter(year == "2018") %>% 
  pull(tempo)

y = tempo_data %>% 
  filter(year == "2019") %>% 
  pull(tempo)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2018 = stats$estimate[1],
  mean_2019 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean tempo for the top streamed songs of 2018 is different than the mean tempo for the top streamed songs of 2019 (p = 0.458, unpaired t-test, df = 98.278)

Lastly, let's check if there is a difference between the means for 2017 and 2019:

```{r, echo=FALSE}
tempo_data = bind_rows(tempo_data_17, tempo_data_19)

tempo_wide = tempo_data %>% 
  pivot_wider(names_from = year, values_from = tempo)

head(tempo_wide, 5) %>% 
  kable(caption = "Tempo between 2017 and 2019 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```

```{r, echo=FALSE}
ggplot(tempo_data, aes(x = year, y = tempo, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Tempo") +
  ggtitle("Boxplots of Tempos in the Top Streamed Songs of 2017 and 2019") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2018 and the other for 2019

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = tempo_data %>% 
  filter(year == "2017") %>% 
  pull(tempo)

y = tempo_data %>% 
  filter(year == "2019") %>% 
  pull(tempo)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2017 = stats$estimate[1],
  mean_2019 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean tempo for the top streamed songs of 2017 is different than the mean tempo for the top streamed songs of 2019 (p = 0.402, unpaired t-test, df = 95.713)

#### Correlations

Because of the results of our T-Test, we also examined the correlation between rank and tempo for each year to see if there was any relationship between the two.

```{r, echo=FALSE}
# compares rank vs tempo for the three separate years
ggplot(songs_2017, aes(rank, tempo)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Tempo vs. Rank: 2017")

corr_2017 = cor(songs_2017$tempo, songs_2017$rank)

ggplot(songs_2018, aes(rank, tempo)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    ggtitle("Tempo vs. Rank: 2018")

corr_2018 = cor(songs_2018$tempo, songs_2018$rank)

ggplot(songs_2019, aes(rank, tempo)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    ggtitle("Tempo vs. Rank: 2019")

corr_2019 = cor(songs_2019$tempo, songs_2019$rank)
```

For each year the correlations were:

- 2017: $r$ = 0.100
- 2018: $r$ = 0.179
- 2019: $r$ = 0.170

### Did the top streamed songs get happier from 2017 to 2019?

Now, we'll summarize the valence data for all three years. 

```{r, echo=FALSE}
valence_2017 = songs_2017 %>% 
  select(valence) %>% 
  summarize(n = n(),mean = mean(valence), sd = sd(valence), max = max(valence), min = min(valence))

valence_2018 = songs_2018 %>% 
  select(valence) %>% 
  summarize(n = n(),mean = mean(valence), sd = sd(valence), max = max(valence), min = min(valence))

valence_2019 = songs_2019 %>% 
  select(valence) %>% 
  summarize(n = n(),mean = mean(valence), sd = sd(valence), max = max(valence), min = min(valence))

valence = bind_rows(valence_2017, valence_2018, valence_2019)

years = tibble(
  year = c(2017, 2018, 2019)
)

bind_cols(years, valence) %>% 
  kable(caption = "Valence Summary") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```

- $year$ is the year the songs are from
- $n$ is the number of songs
- $mean$ is the mean valence for the given year
- $sd$ is the standard deviation of the songs' valence
- $max$ is the song with the highest valence for the year
- $min$ is the song with the lowest valence for the year

From the table, we can see that the mean actually decreased between all three years.

#### T Tests

To test this, we ran an unpaired T Test (Two Mean Test) of valences between 2017 and 2018:


```{r, echo=FALSE}
# combine tempo for 2017 and 2019
valence_data_17 = songs_2017 %>% 
  select(valence, rank) %>% 
  mutate(year = "2017")

valence_data_18 = songs_2018 %>% 
  select(valence, rank) %>% 
  mutate(year = "2018")

valence_data = bind_rows(valence_data_17, valence_data_18)

valence_wide = valence_data %>% 
  pivot_wider(names_from = year, values_from = valence)

head(valence_wide, 5) %>% 
  kable(caption = "Valence between 2017 and 2018 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```



```{r, echo=FALSE}
ggplot(valence_data, aes(x = year, y = valence, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Valence") +
  ggtitle("Boxplots of Valence in the Top Streamed Songs of 2017 and 2018") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2017 and the other for 2018

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = valence_data %>% 
  filter(year == "2017") %>% 
  pull(valence)

y = valence_data %>% 
  filter(year == "2018") %>% 
  pull(valence)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2017 = stats$estimate[1],
  mean_2018 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean valence for the top streamed songs of 2017 is different than the mean valence for the top streamed songs of 2018 (p = 0.862, unpaired t-test, df = 197.532)

Let's run the same unpaired t-test for 2018 and 2019:


```{r, echo=FALSE}
valence_data_19 = songs_2019 %>% 
  select(valence, rank) %>% 
  mutate(year = "2019")

valence_data = bind_rows(valence_data_18, valence_data_19)

valence_wide = valence_data %>% 
  pivot_wider(names_from = year, values_from = valence)

head(valence_wide, 5) %>% 
  kable(caption = "Valence between 2018 and 2019 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```



```{r, echo=FALSE}
ggplot(valence_data, aes(x = year, y = valence, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Valence") +
  ggtitle("Boxplots of Valence in the Top Streamed Songs of 2018 and 2019") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2018 and the other for 2019

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = valence_data %>% 
  filter(year == "2018") %>% 
  pull(valence)

y = valence_data %>% 
  filter(year == "2019") %>% 
  pull(valence)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2018 = stats$estimate[1],
  mean_2019 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean valence for the top streamed songs of 2018 is different than the mean valence for the top streamed songs of 2019 (p = 0.436, unpaired t-test, df = 92.946)

Lastly, let's check if there is a difference between the means for 2017 and 2019:

```{r, echo=FALSE}
valence_data = bind_rows(valence_data_17, valence_data_19)

valence_wide = valence_data %>% 
  pivot_wider(names_from = year, values_from = valence)

head(valence_wide, 5) %>% 
  kable(caption = "Valence between 2017 and 2019 for Top 5 Songs") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))
```

```{r, echo=FALSE}
ggplot(valence_data, aes(x = year, y = valence, fill = year)) +
  geom_boxplot(coef = Inf, alpha = 0.5) +
  geom_point(position = position_jitter(width=0.3, height=0)) +
  xlab("Year") +
  ylab("Valence") +
  ggtitle("Boxplots of Valence in the Top Streamed Songs of 2017 and 2019") +
  theme_minimal() 

```

##### Model

We have two independent samples of songs, one for 2018 and the other for 2019

- $X_i \sim F_1(\mu_1, \sigma_1), \quad i = 1, \ldots, n_1$    
- $Y_i \sim F_2(\mu_2, \sigma_2), \quad i = 1, \ldots, n_2$   


##### Hypothesis Test

$$
H_0: \mu_1 = \mu_2 \\
H_a: \mu_1 < \mu_2
$$

##### Summary of the T-Test:
```{r, echo=FALSE}
x = valence_data %>% 
  filter(year == "2017") %>% 
  pull(valence)

y = valence_data %>% 
  filter(year == "2019") %>% 
  pull(valence)

stats = t.test(x, y, alternative = "less") 

results = tibble(
  mean_2017 = stats$estimate[1],
  mean_2019 = stats$estimate[2],
  t = stats$statistic,
  df = stats$parameter[1],
  pvalue = stats$p.value
)

results %>% 
  kable(caption = "Unpaired T Test Results") %>% 
  kable_styling(position = "left", full_width = FALSE,
                bootstrap_options = c("striped"))

```

We have no evidence from the data that the mean valence for the top streamed songs of 2017 is different than the mean valence for the top streamed songs of 2019 (p = 0.758, unpaired t-test, df = 	97.029)

#### Correlations

Because of the results of our T-Test, we also examined the correlation between rank and valence for each year to see if there was any relationship between the two.

```{r, echo=FALSE}
# compares rank vs tempo for the three separate years
ggplot(songs_2017, aes(rank, valence)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Valence vs. Rank: 2017")

corr_2017 = cor(songs_2017$valence, songs_2017$rank)

ggplot(songs_2018, aes(rank, tempo)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    ggtitle("Valence vs. Rank: 2018")

corr_2018 = cor(songs_2018$valence, songs_2018$rank)

ggplot(songs_2019, aes(rank, tempo)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
    ggtitle("Valence vs. Rank: 2019")

corr_2019 = cor(songs_2019$valence, songs_2019$rank)
```

For each year the correlations were:

- 2017: $r$ = -0.185
- 2018: $r$ = 0.013
- 2019: $r$ = -0.191


## Discussion

In conclusion, we did not find any statistically significant evidence that the tempo and happiness of songs have increased from 2017 to 2019.

### Potential Shortcomings


#### Data accuracy

- We rely on Spotify for accuracy of the data when it comes to the less measurable aspects of a song 

#### Missing key data

- We are looking at only 50 of the top 100 songs for 2019. We would have had more accurate results if we were able to have the top 100.

#### Sample not representative

- We are looking at years directly following each other. 
- Looking at a broader period of time, or years that are not directly next to each other, may make it more representative of the general music taste for the time period.

#### Key assumptions may not hold

- Data might not be approximately normally distributed
- Music taste may change drastically, leading to variables relating differently to rank
- These 3 years we are looking at may not be representative of the rest of the decade/century so far. 

### Future Direction



## References


